---
title: "Day4"
author: "Bart Roes"
date: "December 4, 2020"
output: html_document
---

[Link to Question](https://adventofcode.com/2020/day/4)


```{r setup}
setwd("~/Continuous learning/AdventOfCode/Day4")
my_data <- read.delim("input.txt", header = F, blank.lines.skip = F)

```

> Count the number of valid passports - those that have all required fields. Treat cid as optional. In your batch file, how many passports are valid?

byr (Birth Year)
iyr (Issue Year)
eyr (Expiration Year)
hgt (Height)
hcl (Hair Color)
ecl (Eye Color)
pid (Passport ID)
cid (Country ID)


```{r}

library(stringr)

line_breaks <- my_data$V1 == ""

which(line_breaks)

my_passports <- vector(mode = "list", length = sum(line_breaks)+1)

# group passport information together in single character string
for (i in 1:length(my_passports)){
  if(i == 1){
    my_passports[[i]] <- my_data[1:(which(line_breaks)[i]-1),] %>% 
      str_c(collapse = " ")
  } else  if(i == length(my_passports)){
    my_passports[[i]] <- my_data[(which(line_breaks)[i-1]+1):length(my_data),] %>% 
      str_c(collapse = " ")
  } else {
    my_passports[[i]] <- my_data[(which(line_breaks)[i-1]+1):(which(line_breaks)[i]-1),] %>% 
      str_c(collapse = " ")
  }
}

# define logic

valid_passports <- function(i, passport_list){
  test_passport <- passport_list[[i]]
  
  str_detect(test_passport, "byr") &
    str_detect(test_passport, "iyr") &
    str_detect(test_passport, "eyr") &
    str_detect(test_passport, "hgt") &
    str_detect(test_passport, "hcl") &
    str_detect(test_passport, "ecl") &
    str_detect(test_passport, "pid")
}

passport_test <-sapply(X = 1:length(my_passports), 
       FUN = valid_passports,
       passport_list = my_passports
) 

sum(passport_test)


```



# Question 2

> Count the number of valid passports - those that have all required fields and valid values. Continue to treat cid as optional. In your batch file, how many passports are valid?

byr (Birth Year) - four digits; at least 1920 and at most 2002.
iyr (Issue Year) - four digits; at least 2010 and at most 2020.
eyr (Expiration Year) - four digits; at least 2020 and at most 2030.
hgt (Height) - a number followed by either cm or in:
- If cm, the number must be at least 150 and at most 193.
- If in, the number must be at least 59 and at most 76.
hcl (Hair Color) - a # followed by exactly six characters 0-9 or a-f.
ecl (Eye Color) - exactly one of: amb blu brn gry grn hzl oth.
pid (Passport ID) - a nine-digit number, including leading zeroes.


```{r}
library(tidyverse)

validate_rules <- function(i, passport_list){
  
  test_passport <- passport_list[[i]]
  
  byr <- between(
    substr(test_passport, 
           str_locate(test_passport, "byr:")[2]+1,
           str_locate(test_passport, "(byr:)[0-9]+")[2]) %>% as.numeric(),
    1920,
    2002)
  
  iyr <-  between(
    substr(test_passport, 
           str_locate(test_passport, "iyr:")[2]+1,
           str_locate(test_passport, "(iyr:)[0-9]+")[2]) %>% as.numeric(),
    2010,
    2020)
  
  eyr <- between(
    substr(test_passport, 
           str_locate(test_passport, "eyr:")[2]+1,
           str_locate(test_passport, "(eyr:)[0-9]+")[2]) %>% as.numeric(),
    2020,
    2030)
  
  hcl <- str_count(
    substr(test_passport, 
           str_locate(test_passport, "hcl:#")[2]+1,
           str_locate(test_passport, "(hcl:#)[0-9 | a-f]{6}")[2]),
    "[0-9 | a-z]") == 6
  
  ecl <- substr(test_passport, 
                str_locate(test_passport, "ecl:")[2]+1,
                str_locate(test_passport, "(ecl:)[a-z]+")[2]) %in%
    c("amb", "blu", "brn", "gry", "grn", "hzl", "oth")
  
  pid <- str_count(
    substr(test_passport, 
           str_locate(test_passport, "pid:")[2]+1,
           str_locate(test_passport, "(pid:)[0-9]+")[2]),
    "[0-9]") == 9
  
  hgt <- ifelse(str_detect(test_passport,"cm"),
                between(substr(test_passport, 
                               str_locate(test_passport, "hgt:")[2]+1,
                               str_locate(test_passport, "(hgt:)[0-9]+")[2]) %>%
                          as.numeric(),
                        150,193),
                between(substr(test_passport, 
                               str_locate(test_passport, "hgt:")[2]+1,
                               str_locate(test_passport, "(hgt:)[0-9]+")[2]) %>%
                          as.numeric(),
                        59,76))
  
  byr & iyr & eyr & hgt & hcl & ecl & pid
}


passport_test <-sapply(X = 1:length(my_passports), 
       FUN = validate_rules,
       passport_list = my_passports
) 

sum(passport_test, na.rm = T)


```

